---
title: "A3: SPL Library Checkouts"
output: html_document
author: Nick Ng
---

### Introduction

An introduction of the data and a description of the trends/books/items you are choosing to analyze (and why!)

  The values I'd like to calculate is the proportion of each usage class checked out over time. As our 
technology develops in this day in age, I'm curious on the usage of digital checkouts and how many people are 
transitioning to electronic methods. I'm also curious on whichh types of materials are used the most. Especially 
because of the pandemic, I'm interested in finding out how each material type changed over time. Literature and
fictions are two of the most favored genre and I'm curious on their checked out amounts over time.


### Summary Information

- What is the change in proportion of usage class items over time?
- Which month had the greatest proportion of digital usage items?
- What is the year with the most amount of audiobook checkouts?
- How did the top 5 material type checkout change over time?
- How did the number of fiction versus literate books change over time?

### The Dataset

- Who collected/published the data?

    The data is provided by the Seattle Public Library and the dataset owner is David Christensen  

- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?

    UsageClass, CheckoutType, MaterialType, CheckoutYear, CheckoutMonth, Checkouts, Title, ISBN, 
    Creator, Subjects, Publisher, and PublicationYear.

- How was the data collected or generated? 

    Collected monthly by title for physical and digital items.

- Why was the data collected?  

    This data was collected to not only keep track of which items were checked out, but also allows
    the library to find trends in their popular/non-popular items.

- What, if any, ethical questions do you need to consider when working with this data?  

    Did the borrowers give us consent to keep track of their checkout items and whether or not they 
    think this information is sensitive or not.

- What are possible limitations or problems with this data?   (at least 200 words)

    Some limitations with this data is the ability to see what people checkout together. For example,
    we're not able to see whether people who checkout fiction only like fictions or they enjoy fiction
    best with history. Understanding correlations can allows the library to arrange their books to where 
    checking out books can be closer together. Since this data only shows how many times a certain item is 
    checked out, we cannot determine this information from it.

    One problem with this dataset is that this dataset only provides the items that are checked out 5 times
    within a month. Those that are checked out 4 or less times is not included in this dataset and can cause 
    this dataset to be unrepresentative of the actual trends. There may be a lot of checked out history books 
    that aren't checked out very often but because of this dataset, we aren't able to see it.

    Another problem with this dataset is that it does not provide us with what changed during the pandemic.
    How accessible was the online library? How did the checking out process work? How many people knew 
    about it? These are all questions that can affect about my people use the library during COVID and affect
    the number of checked out items between physical and digital.

### First Trends Over Time chart

```{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("chart1.R")}
```

We're able to see that Ebooks soared during 2020 and this can be the cause of the pandemic.
People were locked down and so they coverted to electronic books and why regular books
dropped too. We also see that video and sound discs are becoming less common over the year.
Audiobooks also increased over time as well.

### Second Trends Over Time Chart

```{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("chart2.R")}
```

In this stacked bar chart, we're able to see that digital usage took over during 2020. 2020 is 
the start of the pandemic and so this can be the reason why digital usage took over.

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and 
more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, warning = FALSE,  message = FALSE, code = readLines("chart3.R")}

```

In this graph, we're able to see that fiction story books are superiorly more popular
than literate books throughout the years. We do notice that people started to read 
less fiction in 2020, the same year of the start of the pandemic but we cannot determine
the cause of this from this dataset.


